---
title: "Stats_506_PS5"
author: "Annie Yannan Niu"
format: 
  html:
    embed-resources: true
execute:
  error: true
editor: visual
---

# Problem set 5

## Github link

<https://github.com/nynn7747/Stats_506_PS5>

## Problem 1. OOP

### a. Define the `rational` class

Function for GCD

```{r}
library(Rcpp)

cppFunction('
int cpp_gcd(int a, int b) {
    while (b != 0) {
        int temp = b;
        b = a % b;
        a = temp;
    }
    return abs(a);
}
')
cpp_gcd(20, 25)
```

Function for LCM

```{r}
cppFunction('
int cpp_lcm(int a, int b) {
    int x = a;
    int y = b;
    while (y != 0) {
            int temp = y;
            y = x % y;
            x = temp;
        }
    return abs(a * b) / abs(x);
}
')

# Test the functions
cpp_lcm(20, 25)   # Should return 100
```

```{r}
# Define the rational S4 class
setClass("rational",
         slots = list(numerator = "numeric",
                      denominator = "numeric"))

# Constructor 
rational <- function(numerator, denominator = 1) {
    # if (denominator == 0) {
    #     stop("Denominator cannot be zero.")
    # }
    obj <- new("rational", numerator = numerator, denominator = denominator)
    validObject(obj)  # Check validity
    return(obj)
}

# Validator 
setValidity("rational", function(object) {
    if (object@denominator == 0) {
        return("Denominator cannot be zero.")
    }
    TRUE
})

# Show method 
setMethod("show", "rational", function(object) {
    cat(object@numerator, "/", object@denominator, "\n")
})

# Simplify method 
setGeneric("simplify", function(object) standardGeneric("simplify"))
setMethod("simplify", "rational", function(object) {
    a <- abs(object@numerator)
    b <- abs(object@denominator)
    gcd <- cpp_gcd(a, b)
    
    # Simplify by devide gcd
    new("rational", numerator = object@numerator / gcd, denominator = object@denominator / gcd)
})

# Quotient method 
setGeneric("quotient", function(object, digit = 2) standardGeneric("quotient"))
setMethod("quotient", "rational", function(object, digit = 2) {
    if (digit %% 1 != 0) {
        stop("Digits must be a non-negative number.")
    }
    result <- object@numerator / object@denominator
    format_string <- paste0("%.", digit, "f")
    round_r <- sprintf(format_string,round(result, digit))
    print(round_r)
    return(result)
})

# Arithmetic methods 
setMethod("+", signature(e1 = "rational", e2 = "rational"), function(e1, e2) {
    lcm_denom <- cpp_lcm(e1@denominator, e2@denominator)
    new_numerator <- e1@numerator * (lcm_denom / e1@denominator) + e2@numerator * (lcm_denom / e2@denominator)
    simplify(new("rational", numerator = new_numerator, denominator = lcm_denom))
})

setMethod("-", signature(e1 = "rational", e2 = "rational"), function(e1, e2) {
    lcm_denom <- cpp_lcm(e1@denominator, e2@denominator)
    new_numerator <- e1@numerator * (lcm_denom / e1@denominator) - e2@numerator * (lcm_denom / e2@denominator)
    simplify(new("rational", numerator = new_numerator, denominator = lcm_denom))
})

setMethod("*", signature(e1 = "rational", e2 = "rational"), function(e1, e2) {
    simplify(new("rational", numerator = e1@numerator * e2@numerator, denominator = e1@denominator * e2@denominator))
})

setMethod("/", signature(e1 = "rational", e2 = "rational"), function(e1, e2) {
    if (e2@numerator == 0) stop("Cannot divide by zero.")
    simplify(new("rational", numerator = e1@numerator * e2@denominator, denominator = e1@denominator * e2@numerator))
})

```

### b. Use your `rational` class to create three objects:

```{r}
# Create instances of Rational numbers
r1 <- rational(24, 6)
r2 <- rational(7, 230)
r3 <- rational(0, 4)
print(c(r1, r2, r3))
```

```{r}
r1
r3
r1 + r2
r1 - r2
r1 * r2
r1 / r2
r1 + r3
r1 * r3
r2 / r3
quotient(r1)
quotient(r2)
quotient(r2, digit = 3)
quotient(r2, digit = 3.14)
quotient(r2, digit = "avocado")
q2 <- quotient(r2, digit = 3)
q2
quotient(r3)
simplify(r1)
simplify(r2)
simplify(r3)
```

### c. Show that your validator does not allow the creation of rational’s with 0 denominator, and check other malformed input to your constructor.

```{r}
r4 <- rational(24, 0)
```

## **Problem 2 - plotly**

### a. Regenerate your plot which addresses the second question from last time:

Does the distribution of genre of sales across years appear to change?

```{r}
# Import data
art_sales <- read.csv("/Users/nynn/Library/CloudStorage/OneDrive-Umich/Umich course/2024_Fall/Stats 506/Stats_506_PS/Stats_506_PS4/df_for_ml_improved_new_market.csv")

# Load necessary libraries
library(tidyverse)
library(plotly)

# Summarize data by year and genre, including NAs/no info as others
genre_summary <- art_sales %>%
  group_by(year) %>%
  summarise(
    Photography = sum(Genre___Photography, na.rm = TRUE),
    Print = sum(Genre___Print, na.rm = TRUE),
    Sculpture = sum(Genre___Sculpture, na.rm = TRUE),
    Painting = sum(Genre___Painting, na.rm = TRUE),
    Other = sum(Genre___Photography == 0 & Genre___Print == 0 &
                  Genre___Sculpture == 0 & Genre___Painting == 0) 
  )

# Reshape data for Plotly
genre_long <- genre_summary %>%
  pivot_longer(cols = c(Photography, Print, Painting, Sculpture, Other), 
               names_to = "Genre", 
               values_to = "Count") %>%
  group_by(year) %>%
  mutate(Percentage = round(Count / sum(Count) * 100, 1))

# Absolute number plot
absolute_plot <- plot_ly(genre_long, x = ~as.factor(year), y = ~Count, color = ~Genre, type = "bar") %>%
  layout(
    title = "Distribution of Genre Over Years - Absolute Number",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Number of Sales"),
    barmode = "stack",
    legend = list(title = list(text = "Genre"))
  )

absolute_plot

# Percentage plot (stacked to 100%)
percentage_plot <- plot_ly(genre_long, x = ~as.factor(year), y = ~Percentage, color = ~Genre, type = "bar", text = ~Genre) %>%
  layout(
    title = "Distribution of Genre Over Years - Percentage",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Percentage of Sales", ticksuffix = "%"),
    barmode = "stack",
    legend = list(title = list(text = "Genre"))
  )

percentage_plot
```

### b. Generate an interactive plot with **plotly** that can address both of these questions from last time:

Is there a change in the sales price in USD over time?

How does the genre affect the change in sales price over time?

```{r}
# Prepare data with median and IQR
df_yearly_price <- art_sales %>%
  group_by(year) %>%
  summarise(
    median_price_usd = median(price_usd, na.rm = TRUE),
    lower_q = quantile(price_usd, 0.25, na.rm = TRUE),  # 25th percentile
    upper_q = quantile(price_usd, 0.75, na.rm = TRUE)   # 75th percentile
  )

# Create the interactive Plotly plot
plot <- plot_ly(df_yearly_price, x = ~year) %>%
  # Add ribbon for IQR
  add_ribbons(ymin = ~lower_q, ymax = ~upper_q, name = "IQR (25th - 75th Percentile)", 
              fillcolor = 'rgba(173, 216, 230, 0.4)', line = list(color = 'transparent')) %>%
  # Add median line
  add_lines(y = ~median_price_usd, name = "Median Price", line = list(color = 'blue', width = 2)) %>%
  # Add median points
  add_markers(y = ~median_price_usd, name = "Median Price", marker = list(color = 'blue', size = 6)) %>%
  # Layout for titles and styling
  layout(
    title = "Sales Price Distribution Over Time with IQR",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Sales Price (USD)"),
    legend = list(title = list(text = "Legend")),
    hovermode = "x unified"  # Display hover info for all elements together
  )

plot
```

```{r}
# Create a dataset for genre
art_sales_genre <- art_sales %>%
  mutate(Genre = case_when(
    Genre___Photography == 1 ~ "Photography",
    Genre___Print == 1 ~ "Print",
    Genre___Sculpture == 1 ~ "Sculpture",
    Genre___Painting == 1 ~ "Painting",
    TRUE ~ "Other"
  )) %>%
  select(year, price_usd, Genre)

art_sales_genre$Genre <- factor(art_sales_genre$Genre, levels = c("Painting", "Photography", "Print", "Sculpture", "Other"))

# Calculate IQR (Interquartile Range) for each genre and year
genre_price_summary <- art_sales_genre %>%
  group_by(year, Genre) %>%
  summarise(
    median_price = median(price_usd, na.rm = TRUE),
    lower_q = quantile(price_usd, 0.25, na.rm = TRUE),  # 25th percentile
    upper_q = quantile(price_usd, 0.75, na.rm = TRUE)   # 75th percentile
  )

# Generate a list of plotly plots for each genre with IQR ribbons and median lines
plots <- genre_price_summary %>%
  group_by(Genre) %>%
  do(plot = plot_ly(data = ., x = ~year) %>%
       add_ribbons(
         ymin = ~lower_q,
         ymax = ~upper_q,
         fillcolor = ~Genre,
         name = "IQR",
         showlegend = FALSE,
         opacity = 0.5
       ) %>%
       add_lines(
         y = ~median_price,
         color = ~Genre,
         name = ~Genre,
         line = list(width = 3),
         hoverinfo = "text",
         text = ~paste("Year:", year,
                       "<br>Median Price:", round(median_price, 2),
                       "<br>Genre:", Genre)
       ) %>%
       layout(
         title = list(text = ~unique(Genre), xref = "paper"),
         xaxis = list(title = "Year"),
         yaxis = list(title = "Median Sales Price (USD)")
       ))

# Use subplot to create a faceted plot with each genre in a separate facet, similar to ggplot's facet_wrap
final_plot <- subplot(plots$plot, nrows = 2, shareX = TRUE, titleX = TRUE)

# Display the final interactive plot with a shared legend at the top
final_plot %>%
  layout(
    title = "Sales Price Over Time by Genre with IQR",
    legend = list(orientation = "h", x = 0.5, xanchor = "center"),
    margin = list(t = 50)
  )
```

## Problem 3. data.table

### a. Generate a table (which can just be a nicely printed tibble) reporting the mean and median departure delay per airport.

```{r}
# Load necessary libraries
library(data.table)
library(nycflights13)

# Convert to data.table format
flights_dt <- as.data.table(flights)
airports_dt <- as.data.table(airports)

# Departure delays
departure_delays <- flights_dt[
  , .(mean_delay = mean(dep_delay, na.rm = TRUE),
      med_delay = median(dep_delay, na.rm = TRUE),
      numflights = .N), by = origin
][numflights >= 10][order(-mean_delay)]

# Join with airports data to get airport names
departure_delays <- departure_delays[
  airports_dt, on = .(origin = faa), nomatch = 0
][, .(name, mean_delay, med_delay)][order(-mean_delay)]

# Display the result for departure delays
print(departure_delays)
```

### b. Generate a second table (which again can be a nicely printed tibble) reporting the mean and median arrival delay per airport.

```{r}
# Arrival delays
arrival_delays <- flights_dt[
  , .(mean_delay = mean(arr_delay, na.rm = TRUE),
      med_delay = median(arr_delay, na.rm = TRUE),
      numflights = .N), by = dest
][numflights >= 10][order(-mean_delay)]

# Join with airports data to get airport names
arrival_delays <- airports_dt[
  arrival_delays, on = .(faa = dest)][, name := coalesce(name, faa)][, .(name, mean_delay, med_delay)][order(-mean_delay)]

# Display all rows for arrival delays
print(arrival_delays)
```

### c. How many flights did the aircraft model with the fastest average speed take? Produce a tibble with 1 row, and entries for the model, average speed (in MPH) and number of flights.

```{r}
planes_dt <- as.data.table(planes)
fastest_aircraft <- flights_dt[
  planes_dt, 
  on = "tailnum", 
  .(model, mph = distance / (air_time / 60)),  # Calculate speed in join
  nomatch = NULL
][
  , .(avgmph = mean(mph, na.rm = TRUE), nflights = .N), by = model  # Aggregate by model
][order(-avgmph)][1]  # Sort by avgmph and take top row

print(fastest_aircraft)
```
